---
title: "Female Representation on the Big Screen"
author: "Mohamed Mohamoud, Jaya Valji, Angela Gaither, Fatima Dineh, Jainaba Jawara"
Affiliation: "INFO-201: Technical Foundations of Informatics - the Information School - University of Washington"
date: "1/22/2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("dplyr")
```
# Report

### Abstract:
This project examines the (under/mis) representation of female actors in the movie industry. Using the Bechdel Test, we observe the trends between how profitable a movie is depending on its Bechdel Test score, which of the three test requirements have the highest number of failed results, and how Bechdel Test scores have changed between 1950-2021. This topic is important because media plays a substantial role in creating and upholding gender norms that get translated into stereotypes embedded in our daily culture. Once these areas have been studied, we can apply it to larger questions begging critical thinking such as how stereotypes of women in films effect their ability to attain key roles.

### Keywords:

**Representation** 

**Women**

**Bechdel Test**

**Movies**

### Introduction:
 Female Representation on the Big Screen is centered around the idea that female actors are not given equal representation on screen in films. Our aim is to use the Bechdel Test as a way to study the patterns of misrepresentation of females in the movie industry. Created by Alison Bechdel in 1985, the Bechdel Test is still relevant today and measures representation by checking three criteria:

 1. The movie must have at least 2 [named] female characters
 2. Who talk to each other 
 3. About something other than a man. 
 In Hollywood, for example, the majority of A-list actors, directors, and writers are men. This has led to their biases of women            tainting the film industry and affecting broader issues such as unequal pay, key roles, and the reinforcement of stereotypes. We all     have our favourite female actors or even directors, but few people are aware of the lack of credit they receive and the major waste     of potential simply due to Hollywood patriarchy.

### Design Situation:

<<<<<<< HEAD
Design Situation:

  Project Framing:The social issue this project aims to address is sexism, specifically when it comes to women in films. Sexism is portrayed through misrepresentation of women where stereotypes are maintained through continuous casting, directing, and script patterns. The project uses the Bechdel Test as a means of measuring the lack of representation and showing its changes over the decades. The project highlights how sexism is structural, and embedded within society through media and the capitalistic economy. 

  Human Values: The human values connected to this project include equality and equity when it comes to gender representation and treatment. Equality refers to all individuals receiving the same resources in same number, whereas equity refers to when specific circumstances of the individuals are considered and resources are allocated based on need. When it comes to finding a solution to the problem of women's representation, it is important increasingly make opportunities for women the same as men (equality), but also go one step further and think about how these opportunities can be efficiently given by effectively focusing on the interests of women first, not men and women together.   

  Direct and Indirect Stakeholders:
The direct stakeholders are the women in Hollywood who are being taken advantage of in all regards by men in power. From all levels in film-making, women of all skills and backgrounds are unfairly represented both economically (if they do not get prominent roles their pay is naturally lessened) as well as being attached to stereotypes that continue gendered ideas of women. The indirect stakeholders in our research would be movie goers and the public who subconsciously make assumptions about women and perpetuate stereotypes in their mind based on the film media they consume. Ultimately, this project and the issue it is addressing can only benefit society as a whole.

Harms and Benefits:
The benefits of this research will be that we will be able to see the inequality in the movie industry in data. It is easy to say that women are being discriminated against, but now there is data compiled to show hard evidence. Possible harm may come from male co-star frustration or conflict over who will be paid less or more. This harm is unimportant however, and ultimately there are no relavent harms that could come from gender equity.
=======
### Project Framing: 

The problem that is of concern here is the discriminatory actions taken against woman in the movie industry. The key issues on hand here are the pay gaps that has lasted decades and still effects up and coming actress and already established ones. Bringing this information to light will show a clear line between how women are discriminated and the actions needed to be taken to fix it. -->

### Human Values:

The human values that are being represented can be quite clear such as sexism that stems from the overall leadership representation of majority men and the hostile work environment that these woman have to endure since this is their work place. Other human values can be the ability to have equal access to things such as power, resources and opportunities. Being fairly treated and respected allows women to be able to have these things at the same rate as men - but that is not the case in the movie industry.

### Direct and Indirect Stakeholders:

The direct stakeholders are the women in the Hollywood who are being taken advantage of in all regards by men in power. From all levels in film-making, women of all skills and backgrounds are more or less are still treated unfairly. The key motivation for women being treated fairly is the fact that this is their vocation and something they enjoy doing. Women in this industry hold the value of wanting to be respected by fellow filmmakers. The indirect stakeholders in our research would be movie goers who watch their favorite actress or director who don't know that they are being discriminated against and might not appear in the movie they watch. -->

### Harms and Benefits:
The benefits of this research will be that we will be able to see the inequality in the movie industry in data. It's easy to say that someone is being discriminated against, but now we will have the data to show the discrimination. The harm that will come will be that male costars will not be paid more but less if their women costars get paid the same as them. This harm is only an if the movie industry changes the pay gap, but if they don't, then no liability will come.
>>>>>>> 16c95615ed8ec9dd24c607eb8c4ef4d9c0f290d9

### Research Questions:

How profitable is a movie depending on its Bechdel Test score?
Unlike our prediction that the most popular films that did not pass the Bechdel Test would have the higher profit, we found no obvious correlation between profitability and Bechdel Test Score. Many popular films that passed the Bechdel Test were highly profitable and some were also not. For example, Cloudy with a Chance of Meatballs 2 FAILED and had a profit of $78000000 and Despicable Me 2 PASSED with a similar profit of $76000000. 	 

In the table, you can see that the movies that passed the Bechdel test didn't guarantee that they ended up being profitable. One of the movies Dredd 3D, passed the Bechdel test, and the binary test had a domestic gross of 13,414,714. Another movie called Captain Phillips failed the Bechdel and binary tests but had a domestic gross of 107,136,417.


Which of the three test requirements have the highest number of failed results?
<<<<<<< HEAD
The test requirement with the highest number of failed tests was women in the film who simply did not have a speaking role. This was surprising and unfortunate to learn because we had thought women would at least have the chance to talk, but maybe that it would mostly be about men. 
=======

The binary test had the most number of failed tests because it was a pass or fail, and there was no in-between. The other test had different rankings, and it wasn't just two options like the binary test

>>>>>>> 16c95615ed8ec9dd24c607eb8c4ef4d9c0f290d9

How have Bechdel Test scores changed between 1874-2021?

If you look at the table, you can see that from 1874 to 1905, it was mainly zeros, and after 1905 it slowly started to get better. In 2021 you will see a lot of threes, so the Bechdel test is changing from 1874 to 2021.

Application and Critical Thinking: How do stereotypes of women in films affect their ability to attain key roles?
This question is important because of the long standing belief that women cant be the star of movies unless there's a romantic element or that they have to be the ones being saved. This has long standing effects to young women that grow up thinking that it is all they will ever be. This also has effects on how directors and writers are picked as well as how much women are paid compared to their male counterparts.

### The Dataset: 
We used the Bechdel test for our data, and the Bechdel test measures the representation of women in fiction. The Bechdel test has three rules: that at least two women are featured, that these women talk to each other, and that they discuss something other than a man. We found an article that used the Bechdel test in female representation in movies, and they gave the article rates in order to figure out the female representation. The score could be one of three meanings: a rating of zero means the movie does not have at least two female presentations. a rating of one, it means that the film has two female actors, but they don't talk to each other. If it has a rating of two, the movie has two female actors, but they talk about men. If they have a three rating, they have two female actors, and they speak to each other and something other than men. We compared about 937 movies, and in our data, we broke it down by title, imdbid, id, rating, and year. If you go to the ratings, you will see which movies passed the Bechdel test and which movies failed the Bechdel test. One of the movies from 2021, called Black Widow, was based on a female superhero and passed the Bechdel test with a three, another movie, Men in Black: International, was based on two male characters, but they still passed the Bechdel test. Another trend we saw was that there is no evident link between a film's budget or income and its Bechdel score.thus, we came to the conclusion that the Bechdel test is not always the greatest way to assess female representation in films.

The movie set is our secondary source which is 1794 rows long and 15 column wide and covers a variety of topics such as budget and gross box office sales to women satisfactory role playing in movies. The who in this data set would be the women who played roles in these movies or not. Some observations I have made with this data is that the connections between the budget and the women's roles in movies and that is something we can really dive into to find how severe the difference is between the budget of movies with men main actors or women.

### Expected Implications:
Our findings have the expected implication that women are still barred from high positions in the media and entertainment industries. This report could be a useful tool for those learning film production and directing or for the general public to be aware of the biases we still have regarding women. This research also has far-reaching implications beyond the film business. The study of historical patterns of female representation is, in and of itself, a very important and relevent subject of research. Our hypothesis is that our studies on films will provide substantial new insights into just how much women are under-represented in many other areas such as STEM, sports, marketing tactics, etc.

### Limitations:
Some limitations that come with this project are the availability and locations of data, and the continuation of movie sagas. Some movies from our Bechdel data are from countries outside of North America, and thus the differences in gender norms and culture must be acknowledged. However, the values behind our project remain - female characters on-screen can be much more than just a love interest or a shallow, supporting character to a male lead and can increased dynamic representation can positively impact female viewers. Furthermore, movies that continue into sagas - like Twilight - can increase data that may pass or fail the Bechdel test and therefore potentially alter the comprehensive, bigger picture. This could be controlled if sagas were reduced to their title films, though this would also disqualify a saga and our data from representing any female character arcs, development, or decreased Bechdel passability like Pirates of the Caribbean. 

### References: 
Mitchell, Carolee. “Women in Movies - Dataset by Carolee.” Data.world, 30 Nov. 2017, https://data.world/carolee/women-in-movies/workspace/file filename=movies.csv.

Yao, Alison Yuhan. “Bechdel Test: Comparing Female Representation Metrics in Movies.” Towards Data Science, 14 Jan. 2022, https://towardsdatascience.com/bechdel- test-comparing-female-representation-metrics-in-movies-6cbade15010f.

Yao, Alison Yuhan. “Movie Bechdel Test Scores.” Kaggle, 2 Jan. 2022, https://www.kaggle.com/alisonyao/movie-bechdel-test-scores.

Creative Commons. (n.d.). Bechdel Test Movie list. Bechdel Test Movie List. Retrieved February 14, 2022, from https://bechdeltest.com/ 

Seth, R., Tsjeng, Z., Coel, M., Bell, D. V., & Devaney, S. (2020, July 15). Why are films failing the Bechdel test when TV has progressed? British Vogue. Retrieved February 14, 2022, from https://www.vogue.co.uk/arts-and-lifestyle/article/the-bechdel-test 

Viswanathan, G. (2021, July 14). What is the Bechdel test and what is its relevance to today's film industry?, Nightingale. Nightingale. Retrieved February 14, 2022, from https://nightingaledvs.com/what-is-the-bechdel-test-and-what-is-its-relevance-to-todays-film-industry/ 
 
### Summary information
We have the title of the movies, and we choose them by overlapping our two data frames and selecting them by which titles were in both. Once we joined them, we were able to take the Bechdel tests rating and see what that meant for the movie data frame. When the binary failed, the rating was zero, one and two, but when it passed, it was three. 
```{r summary, echo= FALSE, message=FALSE, warning=FALSE}
##setwd("~/Documents/_codeDemo/final-project-starter-momohamud001")
source("../source/rating_plot .R")
source("../source/chart2.R")
source("../source/chart3.R")
source("../source/table.R")
```


### Table
In the table, we have incorporated the decades, clean test scores, number of clean tests, ratings, and number of ratings. the clean test is the results from the Bechdel equality test that has forgone peer review to make sure it is accurate and correct. The table gives a snapshot of what was going on throughout the decades. 
```{r Table, echo = FALSE, message=FALSE, warning=FALSE}

join

```

## Plots

### Chart 1:
=======
Chart 1:

This chart is showing a stacked bar plot of the ratings over the decades. As movies  become more
popular over the decades we see a increase in movie production and we see and increase of the ratings. 

```{r chart 1 graph,echo = FALSE, message=FALSE, warning=FALSE}

rating_year_plot

```


### Chart 2:

This chart shows the amount of gender equality that is present in the movie industry by scoring the 
movies from a range of ok for a fair representation of women, to "men" when the movie was classified to only having 
male roles.

```{r chart 2 ,echo = FALSE, message=FALSE, warning=FALSE}
x

```

### Chart 3:

For this chart, we used the text on the y-axis, and the test represents ten categories. Notalk, notalk-disagree, nowomen-disagree, nowomen, dubious-disagree, men,men-disagree, and dubious means that there aren't enough women representation, so they failed the test. Ok-disagree, and ok have women representation, so they passed the test. 

```{r chart 3 ,echo = FALSE, message=FALSE, warning=FALSE}

a

```

