---
title: "Index"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("dplyr")
```

### Summary information
We have the title of the movies, and we choose them by overlapping our two data frames and selecting them by which titles were in both. Once we joined them, we were able to take the Bechdel tests rating and see what that meant for the movie data frame. When the binary failed, the rating was zero, one and two, but when it passed, it was three. 
```{r summary, echo= FALSE, message=FALSE, warning=FALSE}
##setwd("~/Documents/_codeDemo/final-project-starter-momohamud001")
source("../source/rating_plot .R")
source("../source/chart2.R")
#source("../source/table.R")
```


### Table
In the table, we have incorporated the decades, clean test scores, number of clean tests, ratings, and number of ratings. the clean test is the results from the Bechdel equality test that has forgone peer review to make sure it is accurate and correct. The table gives a snapshot of what was going on throughout the decades. 
```{r Table, echo = FALSE, message=FALSE, warning=FALSE}

#join

```

## Plots

##Chart 1:

This chart is showing a stacked bar plot of the ratings over the decades. As movies  become more
popular over the decades we see a increase in movie production and we see and increase of the ratings. 

##Chart 2:

This chart shows the amount of gender equality that is present in the movie industry by scoring the 
movies from a range of ok for a fair representation of women, to "men" when the movie was classified to only having 
male roles.
```{r chart 1 graph,echo = FALSE, message=FALSE, warning=FALSE}

rating_year_plot

```
```{r chart 2 ,echo = FALSE, message=FALSE, warning=FALSE}

x
```

